---
title: "WRN Stats"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
library(magrittr)
library(plyr)
library(tidyverse)
library(reshape2)

data_file <- '~/CPDS/WRN_manuscript/data/NatureDataSpreadsheet_190218_reformattedforstats.xlsx'
```

# Sheet 1 (Fig 2b)

Using Two way ANOVA separately for each cell line 

```{r}
 df <- readxl::read_xlsx('~/CPDS/WRN_manuscript/data/NatureDataSpreadsheet_190218_reformattedforstats.xlsx',
                        sheet = 1) %>% 
  reshape2::melt(variable.name = 'CL_guide') %>% 
  tidyr::separate(CL_guide, into = c('CL', 'guide'), sep = ' ') %>% 
  mutate(is_WRN = factor(grepl('WRN', guide)))
print(table(df$CL, df$guide))

ddply(df, .(CL), function(cur_df) {
  lm(value ~ is_WRN + guide, cur_df) %>% 
    anova() %>% 
    broom::tidy() 
}) %>% filter(term == 'is_WRN') 

ggplot(df, aes(guide, value, group = guide)) + ggbeeswarm::geom_beeswarm() +
  geom_violin() + 
  facet_wrap(~CL)+
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

# Sheet 2 (Fig 2d)

Using Two way ANOVA separately for each cell line 

```{r}
 df <- readxl::read_xlsx('~/CPDS/WRN_manuscript/data/NatureDataSpreadsheet_190218_reformattedforstats.xlsx',
                        sheet = 2) %>% 
  reshape2::melt(variable.name = 'CL_guide') %>% 
  tidyr::separate(CL_guide, into = c('CL', 'guide'), sep = ' ') %>% 
  mutate(is_WRN = factor(grepl('WRN', guide)))
print(table(df$CL, df$guide))

ddply(df, .(CL), function(cur_df) {
  lm(value ~ is_WRN + guide, cur_df) %>% 
    anova() %>% 
    broom::tidy() 
}) %>% filter(term == 'is_WRN') 

ggplot(df, aes(guide, value, group = guide)) + ggbeeswarm::geom_beeswarm() +
  geom_violin() + 
  facet_wrap(~CL)+
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

# Sheet 3 (Fig 3a)

Using Two way ANOVA separately for each cell line 

```{r}
 df <- readxl::read_xlsx('~/CPDS/WRN_manuscript/data/NatureDataSpreadsheet_190218_reformattedforstats.xlsx',
                        sheet = 3) %>% 
  reshape2::melt(variable.name = 'CL_guide') %>% 
  tidyr::separate(CL_guide, into = c('CL', 'guide'), sep = ' ') %>% 
  mutate(is_WRN = factor(grepl('WRN', guide)))
print(table(df$CL, df$guide))

ddply(df, .(CL), function(cur_df) {
  lm(value ~ is_WRN + guide, cur_df) %>% 
    anova() %>% 
    broom::tidy() 
}) %>% filter(term == 'is_WRN') 

ggplot(df, aes(guide, value, group = guide)) + ggbeeswarm::geom_beeswarm() +
  geom_violin() + 
  facet_wrap(~CL)+
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

# Sheet 4 (Fig 3b)

Using Two way ANOVA separately for each cell line 

```{r}
 df <- readxl::read_xlsx('~/CPDS/WRN_manuscript/data/NatureDataSpreadsheet_190218_reformattedforstats.xlsx',
                        sheet = 4) %>% 
  reshape2::melt(variable.name = 'CL_guide') %>% 
  tidyr::separate(CL_guide, into = c('CL', 'guide'), sep = ' ') %>% 
  mutate(is_WRN = factor(grepl('WRN', guide)))
print(table(df$CL, df$guide))

ddply(df, .(CL), function(cur_df) {
  lm(value ~ is_WRN + guide, cur_df) %>% 
    anova() %>% 
    broom::tidy() 
}) %>% filter(term == 'is_WRN') 

ggplot(df, aes(guide, value, group = guide)) + ggbeeswarm::geom_beeswarm() +
  geom_violin() + 
  facet_wrap(~CL)+
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

# Sheet 5 (Fig 4j)

Using Two way ANOVA separately for each cell line 

```{r}
 df <- readxl::read_xlsx('~/CPDS/WRN_manuscript/data/NatureDataSpreadsheet_190218_reformattedforstats.xlsx',
                        sheet = 5) %>% 
  reshape2::melt(variable.name = 'CL_guide') %>% 
  tidyr::separate(CL_guide, into = c('CL', 'guide'), sep = ' ') %>% 
  mutate(is_WRN = factor(grepl('WRN', guide)))
print(table(df$CL, df$guide))

ddply(df, .(CL), function(cur_df) {
  lm(value ~ is_WRN + guide, cur_df) %>% 
    anova() %>% 
    broom::tidy() 
}) %>% filter(term == 'is_WRN') 

ggplot(df, aes(guide, value, group = guide)) + ggbeeswarm::geom_beeswarm() +
  geom_violin() + 
  facet_wrap(~CL)+
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```


# Sheet 6 (Fig ED 3b)

Two-sample t-test for comparing shRFP vs shPSMD2

Two-way ANOVA for comparing shRFP to shWRN

```{r}
 df <- readxl::read_xlsx('~/CPDS/WRN_manuscript/data/NatureDataSpreadsheet_190218_reformattedforstats.xlsx',
                        sheet = 6) %>% 
  reshape2::melt(variable.name = 'CL_guide') %>% 
  tidyr::separate(CL_guide, into = c('day', 'CL', 'guide'), sep = ' ') %>% 
  mutate(is_WRN = (grepl('WRN', guide)),
         is_PSMD2 = (grepl('PSMD2', guide)))
print(table(df$CL, df$guide, df$day))

print('compare shRFP vs shPSMD2')
ddply(df, .(CL, day), function(cur_df) {
    with(cur_df %>% filter(!is_WRN), t.test(value[is_PSMD2], value[!is_PSMD2])) %>% 
    broom::tidy() 
}) 

print('compare shRFP vs shWRN')
ddply(df, .(CL, day), function(cur_df) {
  lm(value ~ is_WRN + guide, cur_df %>% filter(!is_PSMD2)) %>% 
    anova() %>% 
    broom::tidy() 
}) %>% filter(term == 'is_WRN') 

ggplot(df, aes(guide, value, group = guide)) + ggbeeswarm::geom_beeswarm() +
  geom_violin() + 
  facet_wrap(~CL + day)+
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```


# Sheet 7 (Fig ED 10a)

Two-sample t-tests for comparing individual conditions.

Two-way ANOVA for comparing to the two MLH1 guides

```{r}
 df <- readxl::read_xlsx('~/CPDS/WRN_manuscript/data/NatureDataSpreadsheet_190218_reformattedforstats.xlsx',
                        sheet = 7) %>% 
  reshape2::melt(variable.name = 'CL_guide') %>% 
  dplyr::mutate(is_MLH1 = grepl('MLH', CL_guide))
 
print(table(df$CL_guide))

print('HCT116 vs HCT116 Ch2')
with(df, t.test(value[CL_guide == 'HCT116 parental'],
                value[CL_guide == 'HCT116 Ch2'])) %>% 
  broom::tidy()

print('HCT116 vs HCT116 Ch3+5')
with(df, t.test(value[CL_guide == 'HCT116 parental'],
                value[CL_guide == 'HCT116 Ch3+5'])) %>% 
  broom::tidy()

print('HCT116 Ch3+5 Ch2 vs (HCT116 Ch3+5 sgMLH1-1 AND HCT116 Ch3+5 sgMLH1-2)')
with(df %>% filter(CL_guide == 'HCT116 Ch3+5 Ch2-2' | is_MLH1),
     lm(value ~ is_MLH1 + CL_guide)) %>% 
  anova() %>% 
  broom::tidy()

ggplot(df, aes(CL_guide, value, group = CL_guide)) + ggbeeswarm::geom_beeswarm() +
  geom_violin() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```


# Sheet 8 (Fig ED 10b)

Two-way ANOVAs in each case

```{r}
 df <- readxl::read_xlsx('~/CPDS/WRN_manuscript/data/NatureDataSpreadsheet_190218_reformattedforstats.xlsx',
                        sheet = 8) %>% 
  reshape2::melt(variable.name = 'CL_guide') %>% 
  tidyr::separate(CL_guide, into = c('CL', 'guide'), sep = ' ') %>% 
  dplyr::mutate(is_WRN = grepl('WRN', guide),
                is_MLH1 = grepl('MLH1', CL))
 
print(table(df$CL, df$guide))

cur_CL <- c('HCT116-Ch2+sgCh2-2', 'HCT116-Ch3+5+sgCh2-2')
with(df %>% filter(CL %in% cur_CL),
  lm(value ~ CL + guide)
) %>% 
  anova() %>% 
  broom::tidy()

cur_CL <- c('HCT116-Ch2+sgCh2-2', 'HCT116-Ch3+5+sgMLH1-1')
with(df %>% filter(CL %in% cur_CL),
  lm(value ~ CL + guide)
) %>% 
  anova() %>% 
  broom::tidy()

cur_CL <- c('HCT116-Ch2+sgCh2-2', 'HCT116-Ch3+5+sgMLH1-2')
with(df %>% filter(CL %in% cur_CL),
  lm(value ~ CL + guide)
) %>% 
  anova() %>% 
  broom::tidy()


print('As 3-way ANOVA')
lm(value ~ 0 + is_MLH1 + CL + guide, df) %>% 
  anova() %>% 
  broom::tidy()

ggplot(df, aes(CL, value, fill = guide)) + 
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

# Sheet 9 (Fig ED 10e)

Two-way ANOVAs in each case

```{r}
 df <- readxl::read_xlsx('~/CPDS/WRN_manuscript/data/NatureDataSpreadsheet_190218_reformattedforstats.xlsx',
                        sheet = 9) %>% 
  reshape2::melt(variable.name = 'CL_guide') %>% 
  tidyr::separate(CL_guide, into = c('CL', 'guide'), sep = ' ') %>% 
  dplyr::mutate(is_WRN = grepl('WRN', guide),
                is_MLH1 = grepl('MLH1', CL))
 
print(table(df$CL, df$guide))

cur_CL <- c('HCT116-Ch2+sgCh2-2', 'HCT116-Ch3+5+sgCh2-2')
with(df %>% filter(CL %in% cur_CL),
  lm(value ~ CL + guide)
) %>% 
  anova() %>% 
  broom::tidy()

cur_CL <- c('HCT116-Ch2+sgCh2-2', 'HCT116-Ch3+5+sgMLH1-1')
with(df %>% filter(CL %in% cur_CL),
  lm(value ~ CL + guide)
) %>% 
  anova() %>% 
  broom::tidy()

cur_CL <- c('HCT116-Ch2+sgCh2-2', 'HCT116-Ch3+5+sgMLH1-2')
with(df %>% filter(CL %in% cur_CL),
  lm(value ~ CL + guide)
) %>% 
  anova() %>% 
  broom::tidy()

print('As 3-way ANOVA')
lm(value ~ 0 + is_MLH1 + CL + guide, df) %>% 
  anova() %>% 
  broom::tidy()

ggplot(df, aes(CL, value, fill = guide)) + 
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

